### 2010-12-03 report/report.tex
few more plots
Edits with everyone

@@

\section{Method}
% Describe in reasonable detail the algorithm you are using to address this 
% problem. A pseudo-code description of the algorithm you are using is 
% frequently useful. If it makes sense for your project, trace through a 
% concrete example, showing how your algorithm processes this example. The 
% example should be complex enough to illustrate all of the important aspects 
% of the problem but simple enough to be easily understood. If possible, 
% an intuitively meaningful example is better than one with meaningless 
% symbols.
We implemented and tested 8 different learning strategies:
\begin{itemize}
\item \emph{Uniform random hill climbing}: Creates a random neighbor by
  randomly choosing one parameter to adjust, and changing it completely 
  randomly in UniformStrategy(). Then evaluates this neighbor by running 
  the robot with the newly chosen parameters. If this neighbor results in a
  longer distance walked then the previous best neighbor, we save this
  neighbor as the new best neighbor. We then repeat the whole process
  using the best neighbor as a base.
\item \emph{Gaussian random hill climbing}: Creates a random neighbor by
  changing each parameter randomly based on a normal distribution. Then
  evaluates this neighbor similarly to uniform random hill climbing.
\item \emph{N-dimensional policy gradient descent}: Estimates the policy
  gradient by evaluating \emph{t} randomly generated policies
  near an initial parameter vector. Then computes the average score for
  each parameter and adjusts the base policy according to the estimated
  gradient.
\item \emph{Random}: Randomly generates policies in the range of the motion
representation.
\item \emph{Simplex (Nelder-Mead) Method}: Creates a simplex with 6 vertices and tests the fitness of each vertex. The worst vertex is reflected over the centroid. If the reflected point is better than the second worst point and worse than the best point, then the reflected point replaces the worst. If the reflected point is better than the best point, the simplex is expanded in the direction of the reflected point. The better of the reflected and the expanded point replaces the worst point. If the reflected point is worse than the second worst point, then the simplex is contracted away from the reflected point. If the contracted point is better than the reflected point, the contracted point replaces the worst point. If the contracted point is worse than the reflected point, the entire simplex is shrunk.
\item \emph{Linear regression}: To initialize, chooses five random points
and trains on them using least-squares. Then in a loop, takes a fixed-size
step in the direction of the gradient, and retrains on all points so far.
\item \emph{SVM regression}
\item \emph{Evolutionary Neural Network (HyperNEAT)\cite{2}}
\end{itemize}

@@

\section{System Architecture and Implementation}
% Describe how you implemented your system and how you structured it. 
% This should give an overview of the system, not a detailed 
% documentation of the code. The documentation of the code is part of 
% the code you hand in. You might want to comment on high-level design 
% decisions that you made. Also explain how you obtained your
% data and any pre-processing you did to it.

@@

\begin{itemize}
\item \code{Robot} class: Class wrapper for commanding
  motion of the robot.  The \code{Robot} class takes care of the robot
  initialization, communication with the servos, and timing of the
  runs.  In addition, it prevents the servos from ever being commanded
  to a point outside their normal range (0 - 1023) as well as beyond
  points where limbs would collide with parts of the robot body.  The
  main class function, \code{run}, accepts a motion model (any
  function that takes a time argument and outputs a 9 dimensional
  position) and will run the robot using this motion model, including,
  if desired, smooth interpolation over time for the beginning and end
  of the run.

\item \code{Strategy} and \code{RunManager} class: The user has a choice between seven
  different learning strategies: uniform random hill climbing, Gaussian random
  hill climbing, N-dimensional policy gradient descent, random, 
  simplex, linear regression/prediction, and SVM regression/prediction.

\item \code{Parameterized Motion Model} abstract base class: Several functions exist, but none are within a class framework.

\item \code{SineModel} classes: Commands motors to positions
  based on a sine wave, creating a periodic pattern. Parameters are:
  amplitude, wavelength, scale inner vs outer motors, scale left vs right
  motors, scale back vs front motors. Currently a function, not a class.

\item Other derived motion model classes: Several versions of a SineModel
  exist, each with different parameters, though solely as functions.

\item \code{WiiTrackClient} and \code{WiiTrackServer} classes
 and hardware: A Wii remote tracks the location of the robot through an infrared LED mounted on top of the robot. We used CWiid library which is able to interface with the remote via bluetooth to determine the location of the infrared LED. This information is accessed through some functions we wrote and passed to the program. The program gets the robot's position at the beginning of each run and then again at the end. The program then calculates the net change in position and uses that as the distance walked.

@@

\subsection{Methodology}
% What are the criteria you are using to evaluate your method? What
% specific hypotheses does your experiment test? Describe the 
% experimental methodology that you used. What are the dependent and 
% independent variables? For projects in machine learning, what is 
% the training/test data that was used, and why is it realistic or
% interesting? Exactly what performance data did you collect and how 
% are you presenting and analyzing it? Comparisons to competing methods 
% that address the same problem or to variations of your own algorithm 
% are particularly useful.
\begin{itemize}
\item Each algorithm was run on 3 different initial parameter vectors on the
physical robot. We stop each run after plateauing results (no improvement 
for one third of the policies seen so far). We evaluate each method based on the
amount of improvement seen from the initial parameter vectors, and on the
fastest speeds achieved during runs.
\end{itemize}

\subsection{Results}
% Present the quantitative results of your experiments. Graphical data
% presentation such as graphs and histograms are frequently better 
% than tables. Explain what are the basic differences revealed in the 
% data. If your method is stochastic or you analyze multiple datasets, 
% calculate mean and standard error for every quantitative
% performance result you show based on multiple runs.
\begin{itemize}
\item We have done runs of 
3 different initial parameter vectors for random hill climbing, 
Gaussian random hill climbing, policy gradient descent, random,
simplex, and linear regression. We developed several
gaits that were about 4 times faster than the original hand-coded gait.

@@

\item Based on the three trials, linear regression worked the best, followed by simplex. Gaussian random hill climbing performed about as well as random, but it is unclear whether Gaussian would continue to improve and become much better than random if allowed to run for more iterations. Uniform random hill climbing and policy gradient descent performed similarly.

\item We also coded \code{explore\_dimensions} to collect data for dimension
varying plots. We intend to do further explore this by running
\code{explore\_dimension} around a parameter that produces a medium-good gait.
\end{itemize}

\subsection{Discussion}
% Is your hypothesis supported? What conclusions do the results support
% about the strengths and weaknesses of your method compared to other 
% methods? How can the results be explained in terms of the underlying 
% properties of the algorithm and/or the data.
\begin{itemize}
\item  Since only three trials were tested with each algorithm and no algorithm consistently outperformed the others, there is a large standard error for each algorithm. For this reason, it is unclear if any algorithm is superior to another in this application. More trials with each algorithm would be necessary to reach a definitive ranking of the performance of the algorithms.

@@

\begin{itemize}
\item Different motion representations
\item Evolutionary algorithms/HyperNEAT\cite{2}
\item Learning how to turn
\item Open sourcing code for others to extend
\end{itemize}

\section{Conclusion}
% Briefly summarize the important results and conclusions presented in 
% the paper. What are the most important points illustrated by your 
% work? How will your results improve future research and applications 
% in the area?
Because the fastest learned walk was not
significantly faster than the fastest randomly generated walk, we
conjecture that the motion representation for the robot is more
integral to forward speed than the learning algorithm.

\section{References}
% A numbered list of all books, articles, and web pages you used in your 
% work or that you cited in your report. Each item must be cited from 
% somewhere within the text. When citing a URL, include date visited. 
% When citing a commercial product, list the company.

@@

\begin{itemize}
\item \code{optimize.py}: Determines a strategy to try and runs the robot
  with that strategy.
\item \code{RunManager.py}: Deals with all the details of running the robot,
  including choosing an initial parameter, running the robot multiple times, 
  tracking distance walked, and writing to the log file. Also includes the
  \code{explore\_dimensions} method.
\item \code{Strategy.py}: Contains all the different possible strategies, which
  will be passed as objects in \code{optimize.py}. 
\item \code{Robot.py}: Implements the \code{Robot} class, described in
  \secref{implement}.
\item \code{SineModel.py}: Implements a sine based motion model,
  described in \secref{implement}.
\item \code{Motion.py}: Motion helper functions.
\item \code{WiiTrackServer.py}: Broadcasts the position of the infrared LED>
\item \code{WiiTrackClient.py}: Connects to the WiiTrackServer to get the current position information.
\end{itemize}

>>>>

\edit{write this}

@@

\edit{flesh this out... more details}

\section{Method}
% Describe in reasonable detail the algorithm you are using to address this 
% problem. A pseudo-code description of the algorithm you are using is 
% frequently useful. If it makes sense for your project, trace through a 
% concrete example, showing how your algorithm processes this example. The 
% example should be complex enough to illustrate all of the important aspects 
% of the problem but simple enough to be easily understood. If possible, 
% an intuitively meaningful example is better than one with meaningless 
% symbols.

\edit{also describe parameterized motion model}

We implemented and tested 8 different learning strategies:
\begin{itemize}
\item \emph{Uniform random hill climbing}: Creates a random neighbor by
  randomly choosing one parameter to adjust, and changing it completely 
  randomly in UniformStrategy(). Then evaluates this neighbor by running 
  the robot with the newly chosen parameters. If this neighbor results in a
  longer distance walked then the previous best neighbor, we save this
  neighbor as the new best neighbor. We then repeat the whole process
  using the best neighbor as a base.
\item \emph{Gaussian random hill climbing}: Creates a random neighbor by
  changing each parameter randomly based on a normal distribution. Then
  evaluates this neighbor similarly to uniform random hill climbing.
\item \emph{N-dimensional policy gradient descent}: Estimates the policy
  gradient by evaluating \emph{t} randomly generated policies
  near an initial parameter vector. Then computes the average score for
  each parameter and adjusts the base policy according to the estimated
  gradient.
\item \emph{Random}: Randomly generates policies in the range of the motion
representation.
\item \emph{Simplex (Nelder-Mead) Method}: Creates a simplex with 6 vertices and tests the fitness of each vertex. The worst vertex is reflected over the centroid. If the reflected point is better than the second worst point and worse than the best point, then the reflected point replaces the worst. If the reflected point is better than the best point, the simplex is expanded in the direction of the reflected point. The better of the reflected and the expanded point replaces the worst point. If the reflected point is worse than the second worst point, then the simplex is contracted away from the reflected point. If the contracted point is better than the reflected point, the contracted point replaces the worst point. If the contracted point is worse than the reflected point, the entire simplex is shrunk \cite{NMWebsite}.
\item \emph{Linear regression}: To initialize, chooses five random points
and trains on them using least-squares. Then in a loop, takes a fixed-size
step in the direction of the gradient, and retrains on all points so far.
\item \emph{SVM regression}
\item \emph{Evolutionary Neural Network (HyperNEAT)\cite{2}}
\end{itemize}

@@

\edit{Actually write this...}

\section{System Architecture and Implementation}
\seclabel{implement}
% Describe how you implemented your system and how you structured it. 
% This should give an overview of the system, not a detailed 
% documentation of the code. The documentation of the code is part of 
% the code you hand in. You might want to comment on high-level design 
% decisions that you made. Also explain how you obtained your
% data and any pre-processing you did to it.

@@

\begin{itemize}

\item \code{Robot} class: Class wrapper for commanding
  motion of the robot.  The \code{Robot} class takes care of the robot
  initialization, communication with the servos, and timing of the
  runs.  In addition, it prevents the servos from ever being commanded
  to a point outside their normal range (0 - 1023) as well as beyond
  points where limbs would collide with parts of the robot body.  The
  main class function, \code{run}, accepts a motion model (any
  function that takes a time argument and outputs a 9 dimensional
  position) and will run the robot using this motion model, including,
  if desired, smooth interpolation over time for the beginning and end
  of the run.

\item \code{RunManager} class \edit{write this}

\item \code{Strategy} class: The user has a choice between seven
  different learning strategies: uniform random hill climbing, Gaussian random
  hill climbing, N-dimensional policy gradient descent, random, 
  simplex, linear regression/prediction, and SVM regression/prediction.

\item \code{MotionModel} abstract base class: Several functions exist,
  but none are within a class framework. ... ??? ... \code{SineModel}
  classes: Commands motors to positions based on a sine wave, creating
  a periodic pattern. Parameters are: amplitude, wavelength, scale
  inner vs outer motors, scale left vs right motors, scale back vs
  front motors. Currently a function, not a class. ... ??? ... THere
  are also other motion models... Other derived motion model classes:
  Several versions of a SineModel exist, each with different
  parameters, though solely as functions.

\item \code{WiiTrackClient} and \code{WiiTrackServer} classes and
  hardware: A Wii remote tracks the location of the robot through an
  infrared LED mounted on top of the robot. We used the CWiid library
  which is able to interface with the remote via bluetooth to
  determine the location of the infrared LED. This information is
  accessed through some functions we wrote and passed to the
  program. The program gets the robot's position at the beginning of
  each run and then again at the end. The program then calculates the
  net change in position and uses that as the distance
  walked. \edit{Jason: add details}.

@@

\subsection{Methodology}
% What are the criteria you are using to evaluate your method? What
% specific hypotheses does your experiment test? Describe the 
% experimental methodology that you used. What are the dependent and 
% independent variables? For projects in machine learning, what is 
% the training/test data that was used, and why is it realistic or
% interesting? Exactly what performance data did you collect and how 
% are you presenting and analyzing it? Comparisons to competing methods 
% that address the same problem or to variations of your own algorithm 
% are particularly useful.

\editbox{mention using speed as evaluating factor, ``we started the
  robot, and it would collect data on its own.  if it walks otusdie fo
  the viewable area, it tells us and we move it back inside and
  restart it (but the data collection continues)''}

We stop each run after plateauing results (no improvement 
for one third of the policies seen so far)  (WHY??)

Each algorithm was run on 3 different initial parameter vectors on the
physical robot.  ``We decided to evaluate all methods starting at the
same three vectors, to have more of a fair comparison...'' We evaluate
each method based on the amount of improvement seen from the initial
parameter vectors, and on the fastest speeds achieved during runs.

Compare to Juan's run too.

\subsection{Results}
% Present the quantitative results of your experiments. Graphical data
% presentation such as graphs and histograms are frequently better 
% than tables. Explain what are the basic differences revealed in the 
% data. If your method is stochastic or you analyze multiple datasets, 
% calculate mean and standard error for every quantitative
% performance result you show based on multiple runs.

We have done runs of 
3 different initial parameter vectors for random hill climbing, 
Gaussian random hill climbing, policy gradient descent, random,
simplex, and linear regression. We developed several
gaits that were about 4 times faster than the original hand-coded gait.

\editbox{Pull out vectors A, B, and C into separte bulletted list, and
  make table, like below}

\begin{tabular}{|r|c|c|c|}
\hline
                       & A     & B     & C    \\
\hline
Solution depth ???        & 14    & 54    & 21   \\
\hline
Nodes visited          & 992   & 2405  & 383  \\
\hline
Time to find solution  & 139ms & 451ms & 54ms \\
\hline
\end{tabular}

\begin{itemize}

@@

Based on the three trials, linear regression worked the best, followed by simplex. Gaussian random hill climbing performed about as well as random, but it is unclear whether Gaussian would continue to improve and become much better than random if allowed to run for more iterations. Uniform random hill climbing and policy gradient descent performed similarly.

We also coded \code{explore\_dimensions} to collect data for dimension
varying plots. We intend to do further explore this by running
\code{explore\_dimension} around a parameter that produces a medium-good gait.

\edit{fix captions in this section too!}

\figvarp{vectorA}{.65}{Vector A.... more here....}{}

\figvarp{vectorB}{.65}{Vector B.... more here....}{}

\figvarp{vectorC}{.65}{Vector C.... more here....}{}

\figvarp{std_error}{1}{Standard error....}{}

\edit{Jason: add SVM and HyperNEAT results here!}

\subsection{Discussion}
% Is your hypothesis supported? What conclusions do the results support
% about the strengths and weaknesses of your method compared to other 
% methods? How can the results be explained in terms of the underlying 
% properties of the algorithm and/or the data.

\edit{important points... talk about these more :)}

\begin{itemize}
\item  Since only three trials were tested with each algorithm and no algorithm consistently outperformed the others, there is a large standard error for each algorithm. For this reason, it is unclear if any algorithm is superior to another in this application. More trials with each algorithm would be necessary to reach a definitive ranking of the performance of the algorithms.

@@

\edit{Intro sentence here...}

\begin{itemize}
\item More runs and/or longer runs
\item Different motion representations
\item Evolutionary algorithms/HyperNEAT\cite{2}
\item Learning how to turn
\item Open sourcing code for others to extend
\end{itemize}

\section{Conclusion}
% Briefly summarize the important results and conclusions presented in 
% the paper. What are the most important points illustrated by your 
% work? How will your results improve future research and applications 
% in the area?

\edit{make this more awesome}

Because the fastest learned walk was not
significantly faster than the fastest randomly generated walk, we
conjecture that the motion representation for the robot is more
integral to forward speed than the learning algorithm.

@@

\begin{itemize}
\item \code{optimize.py}: Determines a strategy to try and runs the robot
  with that strategy.
\item \code{RunManager.py}: Deals with all the details of running the robot,
  including choosing an initial parameter, running the robot multiple times, 
  tracking distance walked, and writing to the log file. Also includes the
  \code{explore\_dimensions} method.
\item \code{Strategy.py}: Contains all the different possible strategies, which
  will be passed as objects in \code{optimize.py}. 
\item \code{Robot.py}: Implements the \code{Robot} class, described in
  \secref{implement}.
\item \code{SineModel.py}: Implements a sine based motion model,
  described in \secref{implement}.
\item \code{Motion.py}: Motion helper functions.
\item \code{WiiTrackServer.py}: Broadcasts the position of the infrared LED.
\item \code{WiiTrackClient.py}: Connects to the WiiTrackServer to get the current position information.
\end{itemize}
